---
title: "WR-674 Data Issues in Hydrology (but really environmental science generally)"
author:
- "Dr. Matthew Ross"
- 'matt.ross(at)colostate.edu'
- 'Class: T,Th 11-12:15 @ NR 250'
- 'Office Hours: M 3-4, F 2-3 @ NESB B258'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook: default
  #bookdown::pdf_book: default
editor_options: 
  chunk_output_type: console
---

<head> 
<!-- Global site tag (gtag.js) - Google Analytics -->
<script async src="https://www.googletagmanager.com/gtag/js?id=UA-148339061-1"></script>
<script>
  window.dataLayer = window.dataLayer || [];
  function gtag(){dataLayer.push(arguments);}
  gtag('js', new Date());

  gtag('config', 'UA-148339061-1');
</script>
</head>


# (PART) Syllabus {-}

# Environmental Data Science 


This working repository hosts all lecture material, code, assignments, and the
syllabus for Data Issues in Hydrology (WR674) taught at Colorado State University.


## Course Goals

The broadest goal for this course is to make analysis easier and more intuitive
for you. In my experience, code fluency can dramatically increase the efficiency
of your work and significantly add to the the amount
of time you enjoy working whether it be in graduate school, as a post-doc, in government
work, or really anything else. Further, coding is a versatile skill that is transferable
across languages and problems.

Specifically in this course you will learn:

- How to use [R](https://www.r-project.org/), 
[RStudio](https://www.rstudio.com/), 
and the [tidyverse](https://www.tidyverse.org/)

- How to write dynamic scientific documents using 
[RMarkdown](https://rmarkdown.rstudio.com/)

- How to collaboratively code in an open and reproducible framework 
using [Git](https://git-scm.com/) and [GitHub](https://github.com/)

- How to quickly and efficiently organize, clean, and visualize complex 
environmental datasets using tidydata principles

- How to pull down publicly available datasets within R (no more pointing
and clicking on the interwebs)

- How to do basic [geospatial analyses in R](https://geocompr.robinlovelace.net/)

- How to create interactive visualizations using html and JavaScript libraries

- Finally, you will be doing all of this work with your own data and/or
data that is similar to your core datasets. Ideally, the final for this project
will be an analysis that goes into your undergrad, masters, or PhD thesis. 




# Approach

Programming languages are just that, languages. As such, the best way to learn R
or any language is to fully commit to the language and do everything you can
to see it most days of the week for at least half an hour (ideally more). This
can be as easy as trying to write all your homeworks for other classes in R and/or
RMarkdown. Or, it can be as committed as writing everything exclusively in R: 
emails, your personal website, shopping lists, etc... The idea here is immersion 
into the language, ways of asking for help, the community, and more! I promise this
immersive approach will help in the short and long run. This is primarily a coding class,
and any statistical work we do will be guided by you, not me. Later if you
take a stats class, hopefully the coding part will be a breeze!

## Class policies

### Inclusive classroom

To encourage this embedded approach, we will have an inclusive classroom: 
respecting and valuing the diverse backgrounds, perspectives, and identities
of individuals in the class. Students are expected to have an awareness of and
sensitivity to language or actions that may be exclusionary or alienating. I 
am committed to fostering a collaborative and inclusive class and appreciate
any feedback on how to improve my own practice.  

This is particularly important in a class that will have group work,
peer grading, and generally serve individuals with a large range in R fluency. 
We all get better if we work together. Some of the best R training available comes
from groups with explicit inclusivity goals, and I think that is what makes this 
language and community great. I encourage you to check out [RLadies](https://rladies.org/about-us/).
A final note on treatment of others: one of the most important things when 
interacting with others who have more or less coding experience than you is to
learn how to ask effective questions that show you understand and value 
other people's time. We will learn explicitly how to do this in the R 
community, but it is also a good general principle to keep in mind. "What work
can I do to prepare my question in a way that makes it easiest for someone to 
help me?" I want you to ask questions and work collaboratively and do so in a way
that helps you and everyone else.

### Flipped class

Finally, the majority of the class will be "flipped," with lectures delivered
online before class and class time devoted to coding. Learning to code requires
coding and we will try to spend the majority of class time doing assignments,
group work, and live coding problem sets. Before class you will be expected to 
watch short lectures on YouTube that will be posted at least 4 days before class.
These videos will generally be less than 20 minutes and geared directly towards
the material we will work on in class. It is vital that you do this work before
class and I will think of ways to check if needed (I hope not in a grad class).

### Assignments

All assignments will be distributed through GitHub

### Asking for help and code on the internet:

One of the most common ways people learn to code is to use the age-old 
(okay, maybe 10 years) technique of asking the internet for help. This is a great idea!
But! When you do you should ask for help in specific ways that enable people
to answer your question more easily and clearly. Great instructions are here: 
https://blog.revolutionanalytics.com/2014/01/how-to-ask-for-r-help.html
and a more general guide to help in R is here:
https://www.r-project.org/help.html.

### Using other people's code

Inevitably, you will want to perform a task in R that someone else has already 
done. This is useful and part of why the R community is so great. But! If 
you use other people's code, you should:

  a) explitly cite where you got the code or the inspiration for it
  
  b) work hard to understand the code and what it does, break it down into 
  pieces, and try to rebuild it 
  
Code that you did not write and is not cited will be treated as academic 
plagiarism. Assignments will vary in how much code is allowed to be shared 
between students, but the general rule is that you should be submitting your own
code or code from your own team. As an example of best practices, a lot of the
inspiration for this page (and the class in general) comes from Mine Cetikaya-Rundel
and her amazing [STA 199](http://www2.stat.duke.edu/courses/Spring18/Sta199/policies/) 
course at Duke

## Academic integrity

Academic integrity: You are responsible for adhering to all university policies on academic
integrity and student conduct https://tilt.colostate.edu/integrity/knowTheCode/. TILT
has a number of resources for students related to writing and study skills:
https://tilt.colostate.edu/integrity/resourcesStudents/.

## Attendance

This class will heavily depend on you being present.
There will be a participation grade that will reflect a combination of 
your consistent presence, focus in class, and participation in group work. 


## Grading 

```{r, echo=F,message=F,warning=F}
library(knitr)
library(kableExtra)
library(tidyverse)
library(lubridate)

grading <- tibble(Assignments='60%',
                  Final='30%',
                  Participation='10%')

kable(grading)
```

## Schedule 

This schedule will likely change considerably as we move through the course,
but I will keep it updated. Every Thursday there will be an assignment due the
next Friday (of the following week), unless otherwise stated.

```{r, echo=F}



dates <- seq(ymd('2019-08-27'),ymd('2019-12-13'),by=7) %>%
  c(.,seq(ymd('2019-08-29'),ymd('2019-12-13'),by=7)) %>%
  sort(.)


content = c('R, RStudio, Packages, Data Types, and Functions',
            'Projects, Organization, Git, GitHub',
            'Manipulating Data',
            'More Tidying Data ',
            'Generating Dynamic Reports and asking for help',
            'Writing Papers in Rmarkdown (w/ Latex)',
            'Spatial Data in R (Simple Features)',
            'Mixing spatial and data analyses',
            'Raster data',
            'Raster and point data',
            'Visualizing Data (static)',
            'Visualizing Data (dynamic)',
            'Writing functions',
            'Iterating, nesting, purrring',
            'Real data analysis example 1',
            'Real data analysis example 1',
            'Dynamic data retrieval',
            'Web scraping',
            'Intro to Shiny',
            'Shiny as data exploration tool',
            'Watershed analyses',
            'Advanced visualizations (videos and reactive graphs)',
            'Real data analysis example 2',
            'Real data analysis example 2',
            'Bigg(ish) data in R and dtplyr',
            'Parallel processing in R',
            'Fall Break',
            'Fall Break, give thanks',
            'Final project work day',
            'Final project work day',
            'Final presentations',
            'Final presentaions')

urls <- c('https://rstudio.cloud/learn/primers/1',
          'https://happygitwithr.com/',
          'https://rstudio.cloud/learn/primers/2',
          'https://www.jstatsoft.org/article/view/v059i10',
          'https://rmarkdown.rstudio.com/lesson-1.html',
          'https://bookdown.org/yihui/rmarkdown/pdf-document.html',
          'https://geocompr.robinlovelace.net/spatial-class.html#vector-data',
          'https://geocompr.robinlovelace.net/spatial-operations.html') %>%
  c(.,rep(NA,length(dates)-length(.)))


pre_work <- c('Viz and Programming Primers',
              'Happy Git With R',
              'Working with Data Primer',
              'Tidy data',
              'RMarkdown Intro',
              'RMarkdown PDF',
              'Geocomputation with R',
              'Spatial Operations') %>%
  c(.,rep('',length(dates)-length(.)))

pre_videos <- c('Installing R and RStudio',
                'Connecting Git, GitHub, & RStudio',
                'Data manipulation in a script',
                'More complex data',
                'Converting to Rmd',
                'Installing and Using Latex',
                'Intro to Spatial Analysis with Lagos',
                'Spatiala and Data Analyses') %>%
  c(.,rep('',length(dates)-length(.)))

video_urls <- c('https://youtu.be/RN33nzDrWgA',
                'https://youtu.be/Y4sOTFV3FAM',
                'https://youtu.be/Y_Xn8GkDOMw',
                'https://youtu.be/NJ_qSH1Buu0',
                'https://youtu.be/47h2aMx8sY8',
                'https://youtu.be/5WgxUF7B0aY',
                'https://youtu.be/KxNW72CS1Ys',
                'https://youtu.be/MWgratEfvS4') %>%
  c(.,rep(NA,length(dates)-length(.)))


script_urls <- c(NA,NA,'https://github.com/csu-environmental-data-science-2019/Hayman_Fire_NDVI_Trace',
                 'https://github.com/csu-environmental-data-science-2019/Hayman_Fire_NDVI_Trace',
                 '2_ndsi_q.html',
                 'https://github.com/csu-environmental-data-science-2019/LAGOS_Class_Analysis/blob/master/1_beginning_spatial_analysis.Rmd',
                 'https://github.com/csu-environmental-data-science-2019/LAGOS_Class_Analysis/blob/master/2_lake_wq_analysis.Rmd') %>%
  c(.,rep(NA,length(dates)-length(.)))

script <- c('','','Hayman','Hayman again','Q vs Snow cover','Lake analysis','Spatiodata analyses') %>%
  c(.,rep('',length(dates)-length(.)))

data_urls <- c(NA,NA,'https://drive.google.com/file/d/1q9kdnOdnEvZiQS-_lK0yIVJWw_EprGiN/view?usp=sharing',
               'https://drive.google.com/drive/folders/1gETx9YB-6Mg0MWrJwutUN48iNkBXoC0s?usp=sharing',
               NA) %>%
  c(.,rep(NA,length(dates)-length(.)))

data <-  c('','','NDVI','RS Indices','') %>%
  c(.,rep('',length(dates)-length(.)))

schedule <- tibble(`Week`=rep(1:16,each=2),
                   Date = dates,
                   Content=content,
                   `Work Before Class` = pre_work,
                   `Lectures Before Class` = pre_videos) %>%
    mutate(`Work Before Class` = cell_spec(`Work Before Class`,
                                           "html", link = urls),
           `Lectures Before Class` = cell_spec(`Lectures Before Class`,
                                               'html',link=video_urls),
           Scripts = cell_spec(script,'html',link=script_urls),
           Data=cell_spec(data,'html',link=data_urls))

                           


kable(schedule,escape=F) %>%
  kable_styling(bootstrap_options = c("striped", "hover",
                                     "condensed", "responsive")) %>%
  column_spec(3,width='15em')
```

